Estimating Predictive Rate-Distortion via Neural Variational Inference

### 2019-02-15 paper.tex
Updates from Overleaf

\begin{proposition}
Fix $\epsilon$ and $M \in \mathbb{N}$.

	Let $(X_t)_t \sim P$, $(Y_t)_t \sim Q$ be two $M$-th order Markov processes. Assume that
	$$\E_{X_{-\infty}...X_0} \left[\log \frac{P(X_1|X_{-\infty}...X_0)}{Q(X_1|X_{-\infty}...X_0)}\right] < \epsilon$$

	For each point $(R,P$) on the PRD of $(Y_t)_t$, the point $(R+M\epsilon, P - M \epsilon)$ is an upper bound on the PRD of $(X_t)_t$.

\begin{proof}
TODO write
%Storing $S'_0$ and predicting the future on its basis results in $\epsilon H[S_0]$ additional loss in the future.
\end{proof}

@@

\begin{corollary}
If $P, P_n$ are the distributions of $M$-th order Markov processes ($M$ fixed), and if
$$D_{KL}(P_n(\cdot|X_{<0}) || P(\cdot|X_{<0})) \rightarrow 0$$
$$D_{KL}(P(\cdot|X_{<0}) || P_n(\cdot|X_{<0})) \rightarrow 0$$
then the PRD curve of $P_n$ converges to the PRD curve of $P$ uniformly on all of $\mathbb{R}_+$.

@@

\begin{proof}
TODO
\end{proof}

@@

>>>>

%$$\sup_{f : \|f\|_\infty = 1}  \sum_(f(\overrightarrow{X}, X_1))   $$

The total variation distance between two probability measures $P, Q$ on a discrete space is defined as
		$$D_{TV}(P,Q) := \sum_{x} \left|P(x) - Q(x)\right|$$

@@

%We say that a process $P_n$ weakly converges to a process $P$ if for all bounded functions $f$

\begin{corollary}
Assume $P, P_n$ are the distributions of $M$-th order Markov processes ($M$ fixed). If $P_n \rightarrow P$ in the following sense
%$$D_{KL}(P_n(\cdot|X_{<0}) || P(\cdot|X_{<0})) \rightarrow 0$$
%$$D_{KL}(P(\cdot|X_{<0}) || P_n(\cdot|X_{<0})) \rightarrow 0$$
%$$D_{TV}(P, P_n) \rightarrow 0$$
		$$D_{TV}(P_n(X_{1\dots M}), P(X_{1\dots M})) \rightarrow 0$$
and if $P_n(x_{1\dots M}) = 0$ whenever $P(x_{1\dots M}) = 0$,
then the PRD curve of $P_n$ converges to the PRD curve of $P$ uniformly on all of $\mathbb{R}_+$.

@@

\begin{lemma}
Fix $\epsilon$ and $M \in \mathbb{N}$.
Let $(X_t)_t \sim P$, $(Y_t)_t \sim Q$ be two $M$-th order Markov processes. Assume that the two distributions are close in the following sense:
Further, choose $0 < K < +\infty$ such that
	$$- \log P_Y(X_1=x|X_{-\infty}...X_0) \leq K < +\infty$$
	for all $x$ such that $P_X(X_1=x|X_{-\infty}...X_0) > 0$.
For each point $(R,P$) on the PRD of $(Y_t)_t$, the point $(R+\epsilon \cdot \log |S| \cdot M \cdot \exp(K), P - \epsilon \cdot \log |S| \cdot M \cdot \exp(K))$ is an upper bound on the PRD of $(X_t)_t$.
\end{lemma}

\begin{proof}
Let $Z(x_{<0})$ be a code achieving $(P,R)$ on the PRD of $(Y_t)_t$.
When applied to pasts sampled from $(X_t)_t$, it results in rate
\begin{align*}
\sum_{x_{-M...-1}} P_X(x_{-M...-1}) \sum_z p(z|x_{-M...-1}) \log \frac{p(z|x_{-M...-1})}{p_X(z)} \leq \sum_{x_{-M...-1}} P_X(x_{-M...-1}) \sum_z p(z|x_{-M...-1}) \log \frac{p(z|x_{-M...-1})}{p_Y(z)}
\end{align*}
We always know that $- \left(\sum_z p(z|x_{-M...-1}) \log \frac{p(z|x_{-M...-1})}{p_Y(z)}\right) \leq \frac{I[Z,X]}{\exp(-K)} \leq \log |S| \cdot M \cdot \exp(K)$.
Therefore
\begin{align*}
\left|\sum_{x_{-M...-1}} (P_X(x_{-M...-1}) - P_Y(x_{-M...-1})) \sum_z p(z|x_{-M...-1}) \log \frac{p(z|x_{-M...-1})}{p_Y(z)}\right| \leq \epsilon \cdot \log |S| \cdot M \cdot \exp(K)
\end{align*}

A similar argument holds for predictiveness.

%Also, the predictiveness is
%$$D_{KL}(P_Y(

%$$\leq \E_X \log Q_Y(Z) - \E_Y \log Q_Y(Z) $$

%Storing $S'_0$ and predicting the future on its basis results in $\epsilon H[S_0]$ additional loss in the future.
\end{proof}

\begin{proof}

More generally, if $Z$ for a bounded rate amounts to a Markov approximation for bounded length $M$, will still get locally uniform convergence.

Convergence of Markov approximations:

-- code can be computed from context M with high probability

--

@@

Limitation in the case of unbounded Markov order: Process that copies symbol from $M$ words in past. Does this converge to iid Bernoulli weakly, since the finite products of single-variable indicators converge?
\url{https://mathoverflow.net/questions/106234/simple-functions-on-a-product-measure-space}
The PRD curves don't converge.
